---
title: "Lab 1 - Intro to R"
author: "Anyi Guo"
date: "09/10/2018"
output: pdf_document
---

# Vectors
1.Create a vector u that has values -10, -9, -8,. . . ,0. How many different ways can you use?
```{r}
u<- -10:0
u<-seq(-10,0)
```
2.Create another vector v that has values -0.1, 0.4, 0.9, 1.4, . . . , and there are 11 numbers (aka terms) in v. How many different ways can you use?

```{r}
v<-seq(-0.1, by=0.5,length=11)
```
3.Calculate the vector of `u+v` and `u times v`.
```{r}
u+v
u*v
```
4.Increase all terms in u by 1, and then take away 20% from all terms in v.
```{r}
u<-u+1
v<-v*0.8
```
5.Create a vector w that contains all the numbers from u and then v. Assign the length of w to a variable len.
```{r}
w = c(u,v)
```

6.Use a command to return the 14th, 15th and 16th value of w. What about the 2nd, the 5th, 9th and 21st value of w? What is the 23rd value? 
```{r}
w[14:16]
w[c(2,5,9,21)]
w[23]
```

7.Replace the 3rd term of w by 100. Then replace the 7th, 15th and 22nd terms by 200, 300 and 400 simultaneously.
```{r}
w[3]<-100
w[c(7,15,22)]<-c(200,300,400)
```

8.Remove u.
```{r}
rm(u)
```

9.Remove all the objects in the environment.
```{r}
rm(list=ls())
```

## Optional exercises
10.Create a vector p of the values of ex cos(x) at x = 3, 3.1, 3.2, ..., 6.
```{r}
n<-seq(3,6,by=0.1)
p<-exp(n)*cos(n)
```

11.Find the maximum/minimum value in p and the index (position) of that value in p.
```{r}
max(p)
min(p)
mat<-matrix(p)

```
which(mat==max(p))
which(mat==min(p))

12.Sort p in the descending order.
```{r}
sort(p,TRUE)
```

13.Create(4,6,3,4,6,3,. . . ,4,6,3)wherethereare10occurrencesof4.
```{r}
li<- c(4,6,3)
l13<-rep(li,10)
```

14.Create(4,4,. . . ,4,6,6,. . . ,6,3,3,. . . ,3)wherethereare10occurrencesof4,20occurrences of 6 and 30 occurrences of 3.
```{r}
l14<-c(rep(4,10),rep(6,20),rep(3,30))
```

# Matrics
First list is for rows, second list is for columns.
```{r}
a_matrix<-matrix(1:12,nrow=4,dimnames = list(c("one","two","three","four"),c("eins","zwei","drei")))
```

1.Create the following matrix and assign it to the variable b_matrix.
```{r}
b_matrix<-matrix(1:40,nrow= 4,byrow=TRUE)
b_matrix<-b_matrix[,c(1,3,5,7,9)]
dimnames(b_matrix)<-list(c("A","B","C","D"),c("a","b","c","d","e"))
```

2.Extract a sub-matrix from b_matrix named subB as follows. Try to use as many possible ways as you can (positive and negative indices).
```{r}
subB<-b_matrix[c(1,2,4),c(2,3)]
```

3.In R, %*% is an operator for matrix multiplication. Compute a_matrix %*% b_matrix and a_matrix %*% subB. Discuss the results you get from R.
```{r}
#a_matrix %*% b_matrix
a_matrix %*% subB
```

4.Create three vectors x,y,z with integers and each vector has 3 elements. Combine the three vectors to become a 3 × 3 matrix A where each column represents a vector. Change the row names to a,b,c.

```{r}
x<-c(1,2,3)
y<-c(4,5,6)
z<-c(7,8,9)
mat2<-matrix(c(x,y,z),nrow=3)
dimnames(mat2)<-list(c("a","b","c"))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
