---
title: "Lab10_Answer"
author: "Anyi Guo"
date: "04/12/2018"
output: pdf_document
---

```{r}
states<-row.names(USArrests)
states[1:10]
names(USArrests)
```
1) Calculate the mean and variance of each column, by using apply() function.
Hint: `apply(dataset, 1, func)` is to apply the func to each row of dataset, and `apply(dataset, 2, func)` is to apply the func to each column of dataset.
```{r}
# mean
print(apply(USArrests,2,mean))
# variance
print(apply(USArrests,2,var))
```
2) What conclusions can you draw from 1)? And consequently what transformation would you do to your
dataset?
Assault has very high variance compared ot the other variables - we should scale the variables.

3) Perform principal component analysis using the prcomp() function.
```{r}
pr.arrest<-prcomp(USArrests,scale=TRUE)
```
4) Check the results, report the number of PCs and their center, scale, and rotation.
There are 4 PCs.
Center
```{r}
pr.arrest$center
```
Scale
```{r}
pr.arrest$scale
```
Rotation:
```{r}
pr.arrest$rotation
```
5) Plot the first two PCs.

```{r}
biplot(pr.arrest,scale=TRUE)
```

6) What are the standard deviation of each principal component? Based on this result, calculate the variance explained by each PC and the proportion of variance explained by each PC.
```{r}
# standard deviation
pr.arrest$sdev
# variance
pr.arrest$sdev^2
# proportion of variance explained
pr.arrest.var<-pr.arrest$sdev^2
pve<-pr.arrest.var/sum(pr.arrest.var)
pve
```
First PC: 62.0%
Second PC: 24.7%
Third PC:8.9%
Fourth PC: 4.3%

7) Plot the PVE explained by each component as well as the cumulative PVE. Hint: the cumulative PVE can be obtained by the cumsum() function.

For each component:
```{r}
plot(pve,xlab="Principal Component", ylab="Proportion of Variance Explained", type="b",ylim=c(0,1))
```

Cumulative PVE:
```{r}
plot(cumsum(pve),xlab="Principal Component", ylab="Cumulative Proportion of Variance Explained", type="b",ylim=c(0,1))
```
