---
title: "Lab3 Answer"
author: "Anyi Guo"
date: "16/10/2018"
output:
  pdf_document: default
  word_document: default
---
# Lab 3 Distance vs. Fire Damage

1) Use R to estimate the unknown parameters of the hypothesised model. Find the least squares estimates of the slope `b1` and intercept `b0` on the printout. Write down the least squares equation for this model.
```{r}
library(ISLR)
library(MASS)
Distance<-c(3.4,1.8,4.6,2.3,3.1,5.5,0.7,3.0,2.6,4.3,2.1,1.1,6.1,4.8,3.8)
Damage<-c(26.2,17.8,31.3,23.1,27.5,36.0,14.1,22.3,19.6,31.3,24.0,17.3,43.2,36.4,26.1)
FireData<-data.frame(Distance,Damage)
fire_fit<-lm(Damage~Distance,data=FireData)
```

* Least squares estimates of the slope `b1` is:`4.919`  
* Intercept `b0`  is: `10.278`  
* <span style="color:blue">Least squares equation is:</span>
`Damage = 10.278 + 4.919 * Distance`


2) Plot the dataset and the least squares line on the same figure.

```{r}
plot(Distance,Damage)+abline(fire_fit)
```

3) How to interpret the slope `b1` and y-intercept `b0` of the least squares line?

* `b1`: For every increase of 1 mile, the fire damage would cost $4919.3 more. 
* `b0`: Even if the fire takes place exactly where the fire station is (e.g. Distance=0), it would still cost $10,277.9 to put off the fire. 

4) Measuring the extent to which the model fits the data.
```{r}
summary(fire_fit)
```
a. What is the RSE of this model? How to interpret it?

* `RSE` is 2.316. It means that even if this is correct, any prediction on cost of damage on the basis of distance would still be off by $2316 on average


b. What is the R-squared value? How to interpret it?

* `R-squared` is 0.9176 (Using Adjusted R-Squared). It means that 91.76% of the variation of Damage Cost can be explained by changes in Distance.

5) Test the null hypothesis that `b1` is 0. Find the t-value and p-value. What conclusions can you draw from the two values?
* `t-value` is 12.525
* `p-value` is 1.248e-08.
* The low p-value(< 5%) indicates that there is a relationship between Distance and Damage Cost, so we can reject the null hypothesis.

6) Predict the value of damage (y) for a new set of distances (x = 0.5, 1.5, 2.5, 3.5). Can you obtain the above new y's simultaneously?

```{r}
predict(fire_fit,data.frame(Distance=c(0.5,1.5,2.5,3.5)))
```

7) Find the confidence interval for the prediction of damage (y) for a new set of distance (x = 0.5, 3, 5.5). Can you obtain the above confidence intervals simultaneously?

```{r}
predict(fire_fit,data.frame(Distance=c(0.5,3,5.5)),interval="confidence")
```

8) Find the prediction interval for the prediction of damage (y) for a new set of distance (x = 0.5, 3, 5.5). Can you obtain the above confidence intervals simultaneously?
```{r}
predict(fire_fit,data.frame(Distance=c(0.5,3,5.5)),interval="prediction")
```
9) Compare your results in 7) and 8) and comment on the confidence intervals and prediction intervals.

* Althought the fit figures (first column) is the same for both Prediction and Confidence intervals, Prediction intervals have wider lower and upper boundary than Confidence intervals.

10) Plot the confidence intervals and prediction intervals in the same figure.
```{r}
More_Distance<-data.frame(Distance=seq(0,7,length=51))
p_conf<-predict(fire_fit,interval="confidence",newdata=More_Distance)
p_pred<-predict(fire_fit,interval="prediction",newdata=More_Distance)
p<-plot(Distance,Damage,data=FireData)+ abline(fire_fit)+lines(More_Distance$Distance,p_conf[,"lwr"],col="red",type="b",pch="+")+lines(More_Distance$Distance,p_conf[,"upr"],col="red",type="b",pch="+")+lines(More_Distance$Distance,p_pred[,"lwr"],col="blue",type="b",pch="*")+lines(More_Distance$Distance,p_pred[,"upr"],col="blue",type="b",pch="*")

```
