---
title: "Lab4_Answer"
author: "Anyi Guo"
date: "23/10/2018"
output: pdf_document
---
# Lab 4 Logistic Regression
##  Problem Statement

A researcher is interested in how variables, such as GRE (Graduate Record Exam scores), GPA (grade point average) and rank (prestige of the undergraduate institution), affect admission into graduate school. The response variable, admit/do not admit, is a binary variable.

## Dataset
The dataset is included in the package aod. Install the package and include package using the command library(aod).
```{r}
library(aod)
mydata<-read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")
# take a look at the first few rows
head(mydata)
```
Using the following command to load the dataset
##   admit gre  gpa rank
##1 0 380 3.61 3
##2 1 660 3.67 3
##3 1 800 4.00 1
##4 1 640 3.19 4
##5 0 520 2.93 4
##6 1 760 3.00 2

[More on reading and writing CSV files, see here: https://swcarpentry.github.io/r-novice-inflammation/11-supp-read-write-csv/index.html]

This dataset has a binary response (outcome, dependent) variable called admit. There are three predictor variables: gre, gpa and rank. We will treat the variables gre and gpa as continuous. The variable rank takes on the values 1 through 4. Institutions with a rank of 1 have the highest prestige, while those with a rank of 4 have the lowest.

Questions

1) Get basic descriptives for the entire data set using summary(). View the dataset using View().  

2) How many observations are there in this dataset?

3) Get the standard deviations for the first three variables (i.e., admit, gre and gpa). Hint: use sapply to apply the sd function to each variable in the dataset: sapply(mydata, sd). Now get the mean admit, gre and gpa in a similar way.

```{r}
 mydata <- read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv") ## view the first few rows of the data
head(mydata)
1
```

4) Convert rank to a factor to indicate that rank should be treated as a categorical variable. (Hint: use factor() function)
[More on factors, see the tutorial here: https://swcarpentry.github.io/r-novice-inflammation/12-supp-factors/ index.html]
5) Estimate a logistic regression model using the glm function, and get the results using the summary command.
6) Do you notice variable rank is replaced with categorical variables rank2, rank3, and rank4 that can only take values of 0 or 1? Recall that the original variable rank can take values of 1, 2, 3, or 4. Why isn’t a variable rank1 needed? If rank is 1, what are the values of rank2, rank3 and rank4?

7) From the z-statistics and p-values of the variables, report which variables are statistically significant.

8) Use the model to predict the training dataset and store the results to a vector of probabilities admit.prob.

9) Create another vector admit.pred to show 0 or 1 for admit.prob. Let’s set the value to be 0 if the probability is less than 0.5, and 1 if the probability is no less than 0.5.

10) Using table() function to create a confusion matrix to determines how many observations were correctly or incorrectly classified. Calculate the percentage that the observations were correctly classified.

11) Use the model to predict the average cases in each rank, that is, four new data with mean gre, mean gpa and rank from 1 to 4.
2