---
title: "Lesson 1 - Big Data Analytics using R"
author: "Anyi Guo"
date: "01/10/2018"
output: html_document
---

• Create a vector x
– starting from 5.3
– ending at 8.00
– length is 10 

```{r}
x<- seq(5.3,8, length = 10)
```

Create another vector y
– starting from 3.5
– ending at or less than 7.9
– each term is 0.4 more than the previous one
y= 3.5, 3.9, 4.3, …

```{r}
y<-seq(3.5, 7.9, by= 0.4)
```


Slide 16: Creating matrix B
```{r}
data<-matrix(1:40,4,10,byrow=TRUE)
data[,-c(2,4,6,8,10)]
data[c(2,4),c(3,5)]
```
Slide 16: Creating matrix C

```{r}
a<-matrix(c(1),1,1)
a<-matrix(c(1,1),1,2,byrow=TRUE)
```

Use set.seed() to generate random numbers

Use fix() to view data in a spreadsheet like window. The window must be closed before
further R commands can be entered. 

dim() to see the dimension of the dataset

na.omit() to remove rows with missing values
```{r}
Auto = na.omit(Auto)
```

search() to see currently loaded packages

removes element x in the environment
```{r}
rm(x)
```
remove everything loaded in the environment

```{r}
rm(list=ls())
```

lapply(target item, function to apply)
sapply(target item, function to apply)

Difference between lapply() and sapply():
lapply() returns a list, but sapply returns a vector (which is on the same line) which is a simplified version of this list